---
title: "CogSci2026"
output: html_document
---

Supress messages in this document so that the output is cleaner.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

Import packages.

```{r}
rm(list = ls())
options(warn = -1)
library(tidyverse)
library(lme4)
library(lmerTest)
library(ggplot2)
library(aida)
library(BayesFactor)
library(pwr)
library(effsize)
library(HDInterval)
library(tidybayes)
library(purrr)
library(brms)
```

Set up theme for ggplot.

```{r}
theme_set(theme_aida())

##################################################
## CSP-colors
##################################################
CSP_colors = c(
  "#7581B3", "#99C2C2", "#C65353", "#E2BA78", "#5C7457", "#575463",
  "#B0B7D4", "#66A3A3", "#DB9494", "#D49735", "#9BB096", "#D4D3D9",
  "#414C76", "#993333"
  )
# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = CSP_colors)
}
scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = CSP_colors)
}
```

# Exp 1a: Choose and Explain - Recall Task
## Load data, data preparation for recall task
```{r}
data_recall_batch1 <- read.csv("../data+analysis/02-exp2a-choose-explain-recall/01b-processed-data-pilot-2a.csv")
data_recall_batch2 <- read.csv("../data+analysis/02-exp2a-choose-explain-recall/03-raw-data-batch2-cogsci2026_annotated.csv")
```

Rename reproClass in batch1 to responseType, and make values in responseType capitalized first letter.
```{r}
data_recall_batch1 <- data_recall_batch1 %>%
  rename(responseType = reproClass) %>%
  mutate(
    responseType = str_to_sentence(responseType)
  )
```

Stack two dfs together.
```{r}
data_recall <- bind_rows(
  data_recall_batch1,
  data_recall_batch2
)
```

Substract responses into a cleaned df.
```{r}
library(stringr)

data_recall2 <- data_recall %>%
  mutate(
    response     = na_if(str_squish(as.character(response)), ""),
    responseType = na_if(str_squish(as.character(responseType)), "")
  )

data_clean <- data_recall2 %>%
  group_by(submission_id) %>%
  summarise(
    response     = first(na.omit(response)),
    responseType = first(na.omit(responseType)),
    .groups = "drop"
  )
```

Rename Causal in responseType to Causation for consistency.
```{r}
data_clean <- data_clean %>%
  mutate(
    responseType = if_else(responseType == "Causal", "Causation", responseType)
  )
```

Encode responseType into 0 and 1. If causation, then 1, else 0.
```{r}
data_clean <- data_clean %>%
  mutate(
    responseType_bin = if_else(
      responseType == "Causation",
      1L,
      0L
    )
  )
```

## Plot
```{r}
# ----- helper: bootstrap CI for multinomial proportions within each response -----
boot_ci_props <- function(df, B = 2000, conf = 0.95) {
  # df must contain columns: responseType
  types <- sort(unique(df$responseType))
  n <- nrow(df)

  # bootstrap resampling of rows (trials)
  boot_mat <- replicate(B, {
    idx <- sample.int(n, size = n, replace = TRUE)
    tab <- table(factor(df$responseType[idx], levels = types))
    as.numeric(tab) / n
  })

  alpha <- (1 - conf) / 2
  tibble(
    responseType = types,
    prop_hat = as.numeric(table(factor(df$responseType, levels = types))) / n,
    ci_low  = apply(boot_mat, 1, quantile, probs = alpha),
    ci_high = apply(boot_mat, 1, quantile, probs = 1 - alpha),
    n = n
  )
}

# ----- compute plot data with bootstrap CIs -----
data_plot <- data_clean %>%
  filter(!is.na(response), !is.na(responseType)) %>%
  group_by(response) %>%
  group_modify(~ boot_ci_props(.x, B = 2000, conf = 0.95)) %>%
  ungroup() %>%
  rename(proportion = prop_hat)

# ----- compute plot data with bootstrap CIs -----
data_plot %>%   mutate(
    experiment = "experiment1a"
  ) -> data_plot_exp1a

# ----- plot -----
ggplot(data_plot, aes(x = responseType, y = proportion, fill = responseType)) +
  geom_col() +
  geom_errorbar(
    aes(ymin = ci_low, ymax = ci_high),
    width = 0.15
  ) +
  facet_wrap(~ response) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL,                 # omit x-axis title
    y = NULL,
    fill = "Response"
  ) +
  coord_flip() +
  theme_aida() +
  theme(
    legend.position = "top",
    text = element_text(size = 16),
    axis.text = element_text(size = 14),
    strip.text = element_text(size = 16),
    axis.title.y = element_text(size = 16)
  ) -> p_proportion_choice

p_proportion_choice

ggsave(
  filename = "results_exp2a.png",
  plot = p_proportion_choice,
  width = 8,
  height = 5,
  dpi = 300
)
  
```

## Fit a glm model
```{r}
data_clean <- data_clean %>%
  mutate(
    response = relevel(
      factor(response),
      ref = "only xeliherb"
    )
  )

m2_bayes <- brm(
  responseType_bin ~ response + (1 | submission_id),
  data = data_clean,
  family = bernoulli(link = "logit"),
  prior = c(
    prior(normal(0, 1.5), class = "b"),          # fixed effects
    prior(normal(0, 1.5), class = "Intercept"),
    prior(exponential(1), class = "sd")          # random effects SD
  ),
  chains = 4,
  cores = 4,
  iter = 4000,
  warmup = 1000,
  seed = 123,
  backend = "cmdstanr"
)

summary(m2_bayes)

hypothesis(
  m2_bayes,
  "responseboth xeliherb and ralocrop > 0"
)
# responseboth xeliherb and ralocrop    2.235      1.095   2.041   0.0413 * 
```

# Exp 1b: Choose and Explain - Advice Task
## Data preparation
Load the data.

```{r}
data <- read.csv("../data+analysis/02-exp2b-choose-explain-advice/01-data-annotated-pilot-04.csv")
```
Collapse some categories.
"nonsense", "none" -> "none"
"directive-xeliherb", "correlation" -> "neutral"
"directive-ralocrop", "causation" -> "causation"
```{r}
data <- data %>%
  mutate(
    responseType = case_when(
      responseType %in% c("nonsense", "none") ~ "none",
      responseType %in% c("directive-xeliherb", "correlation") ~ "neutral",
      responseType %in% c("directive-ralocrop", "causation") ~ "causation",
      TRUE ~ responseType
    )
  )
```

Encode response type into 0 and 1.
1 are: "causation", "directive-ralocrop"
0 are everything else.

```{r}
library(dplyr)

data_clean <- data %>%
  group_by(submission_id) %>%
  summarise(
    response = response[measure == "choice"],
    responseType = responseType[measure == "reproduction"],
    .groups = "drop"
  )  %>% mutate(
    responseType_bin = if_else(
      responseType %in% c("causation", "directive-ralocrop"),
      1L,
      0L
    )
  )
```

## Plot
```{r}

data_plot <- data_clean %>%
  filter(!is.na(response), !is.na(responseType)) %>%
  group_by(response) %>%
  group_modify(~ boot_ci_props(.x, B = 2000, conf = 0.95)) %>%
  ungroup() %>%
  rename(proportion = prop_hat)

data_plot %>%   mutate(
    responseType = str_to_sentence(responseType)
  ) -> data_plot

# Prepare data for plotting both exps
data_plot_exp1b <- data_plot %>%
  mutate(
    experiment = "experiment1b"
  )

# ----- plot -----
ggplot(data_plot, aes(x = responseType, y = proportion, fill = responseType)) +
  geom_col() +
  geom_errorbar(
    aes(ymin = ci_low, ymax = ci_high),
    width = 0.15
  ) +
  facet_wrap(~ response) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL,                 # omit x-axis title
    y = NULL,
    fill = "Response"
  ) +
  coord_flip() +
  theme_aida() +
  theme(
    legend.position = "top",
    text = element_text(size = 16),
    axis.text = element_text(size = 14),
    strip.text = element_text(size = 16),
    axis.title.y = element_text(size = 16)
  ) -> p_proportion_choice_advice

p_proportion_choice_advice

ggsave(
  filename = "results_exp2b.png",
  plot = p_proportion_choice_advice,
  width = 8,
  height = 5,
  dpi = 300
)
```

## Fit a glm model

```{r}
data_clean <- data_clean %>%
  mutate(
    response = relevel(
      factor(response),
      ref = "only xeliherb"
    )
  )

m2_bayes <- brm(
  responseType_bin ~ response + (1 | submission_id),
  data = data_clean,
  family = bernoulli(link = "logit"),
  prior = c(
    prior(normal(0, 1.5), class = "b"),          # fixed effects
    prior(normal(0, 1.5), class = "Intercept"),
    prior(exponential(1), class = "sd")          # random effects SD
  ),
  chains = 4,
  cores = 4,
  iter = 4000,
  warmup = 1000,
  seed = 123,
  backend = "cmdstanr"
)

summary(m2_bayes)

hypothesis(
  m2_bayes,
  "responseboth xeliherb and ralocrop > 0"
)
```
# Combined plot for both exp 1a and 1b
```{r}
data_plot_combined <- bind_rows(
  data_plot_exp1a,
  data_plot_exp1b
)


data_plot_combined <- data_plot_combined %>%
  mutate(
    response = str_replace_all(
      response,
      c(
        "both xeliherb and ralocrop" = "both xeliherb\nand ralocrop",
        "only xeliherb" = "only\nxeliherb"
      )
    )
  )

CSP_colors_combined_plot = c(
  "#C65353", "#7581B3", "#99C2C2", "#E2BA78", "#5C7457", "#575463",
  "#B0B7D4", "#66A3A3", "#DB9494", "#D49735", "#9BB096", "#D4D3D9",
  "#414C76", "#993333"
  )


ggplot(data_plot_combined, aes(x = responseType, y = proportion, fill = responseType)) +
  geom_col() +
  geom_errorbar(
    aes(ymin = ci_low, ymax = ci_high),
    width = 0.15
  ) +
  facet_grid(rows = vars(response), cols = vars(experiment),
             labeller = labeller(
      experiment = as_labeller(c(
        "experiment1a" = "Experiment 1a: recall",
        "experiment1b" = "Experiment 1b: advice"
      )))) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = CSP_colors_combined_plot) +
  labs(
    x = NULL,                 # omit x-axis title
    y = NULL,
    fill = "Response"
  ) +
  coord_flip() +
  theme_aida() +
  theme(
    legend.position = "none",
    text = element_text(size = 16),
    axis.text = element_text(size = 14),
    strip.text = element_text(size = 14),
    strip.text.y = element_text(size = 14),
    axis.title.y = element_text(size = 16)
  ) -> p_proportion_choice_combined


p_proportion_choice_combined

ggsave(
  filename = "results_exp2_combined.png",
  plot = p_proportion_choice_combined,
  width = 7.2,
  height = 3.6,
  units = "in",
  dpi = 400
)
```

# Exp. 2
## Data preparation

```{r}
data <- read.csv("../data+analysis/03-exp3ab-speakerRelevance/data_preprocessed_allFactors.csv")
```

Define, encode and relevel factors.

```{r}
library(dplyr)
library(tidyr)
# Define factors
data %>% mutate(
  submission_id = as.factor(submission_id),
  informationSource = as.factor(informationSource),
  informationSource = relevel(informationSource, ref = "indirect"),
  scienceTeam = as.factor(scienceTeam),
  scienceTeam = relevel(scienceTeam, ref = "Cultivation"),
) -> data

# Reshape data to long format for decision points
data %>%
  select(decision1, decision2, informationSource, scienceTeam, submission_id) %>%
  pivot_longer(
    cols = c(decision1, decision2),
    names_to = "decision_point",
    values_to = "decision"
  )   %>% mutate(
    decision_point = if_else(decision_point == "decision1", 0L, 1L),
  ) -> data_long

# Covert decisions to [0,1] for zoib likelihood
data_long <- data_long %>%
  mutate(
    decision = decision / 100
  )
```

## Plot
```{r}
library(ggplot2)

# Prepare data for plotting
data_plot <- data %>%
  mutate(
    scienceTeam = factor(
      scienceTeam,
      levels = c("Cultivation", "Localization"),
      labels = c("intervention-oriented", "diagnostic-oriented")
    )
  ) %>%
  mutate(
    scienceTeam = relevel(scienceTeam, ref = "intervention-oriented")
  ) %>%
  mutate(    update = (update + 0.05)/(100 + 0.05))

# Plot mean change of causal confidence by information source and speaker goal orientation
r <- data_plot %>%
  ggplot(aes(x = informationSource, y = update, color = scienceTeam)) +
  stat_summary(
    fun.data = mean_cl_boot,
    geom = "point",
    position = position_dodge(0.5),
    size = 3
  ) +
  stat_summary(
    fun.data = mean_cl_boot,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(0.5),
    linewidth = 0.8
  ) +
  labs(
    x = "Directness of Information",
    y = "Mean Change of Causal Confidence",
    color = "Speaker Goal Orientation"
  ) +
  theme_aida() +
  theme(
    legend.position = "top",
    text = element_text(size = 16),
    axis.text = element_text(size = 14),
    strip.text = element_text(size = 16),
    axis.title.y = element_text(size = 16)
  ) +
    theme(
    text = element_text(size = 16),
    axis.text = element_text(size = 14),
    strip.text = element_text(size = 15),
    strip.text.y = element_text(size = 12, angle = 0),
    plot.margin = margin(8, 8, 8, 14),
    panel.spacing.y = unit(0.25, "lines")
  )


# Display the plot
r

# Save the plot to a file
ggsave("results_exp_3.png", plot = r, width = 7.2, height = 4,
       units = "in",dpi = 400)
```
## Fit model
```{r}
library(brms)
kernel_long_mixed_decision_point <- bf(
  decision ~ decision_point + decision_point:(informationSource * scienceTeam) +
    (1 + decision_point | submission_id)
)
lik <- zero_one_inflated_beta(link = "logit", link_phi = "log")

prior_zoib <- c(
  # mu (mean of the Beta part) on logit scale
  prior(normal(0, 1), class = "b"),
  prior(student_t(3, 0, 2.5), class = "Intercept"),

  # phi (precision) on log scale
  prior(exponential(2), class = "phi"),

  # zoi: probability of being in the {0,1} inflation component
  # (logit scale internally; brms default is beta(1,1) on prob scale)
  prior(beta(8, 2), class = "zoi"),  # mildly prefers high boundary mass

  # coi: conditional probability of 1 given inflation (vs 0)
  prior(beta(2, 2), class = "coi")   # weakly centered at 0.5
)

m_zoib_long_mixed_decision_point <- brm(
  formula = kernel_long_mixed_decision_point,
  data = data_long,
  family = lik,
  prior = prior_zoib,
  chains = 4,
  cores = 4,
  iter = 4000,
  warmup = 1000,
  backend = "cmdstanr",
  seed = 123
)
```
```{r}
summary(m_zoib_long_mixed_decision_point)
```

## Pair-wise comparison
```{r}
# Posterior expected predictions (response scale) for:
#  1) Cell means (2x2 at decision point 2)
#  2) Delta interaction:
#        Delta_Team_byInfo = (direct_Loc - direct_Cult) - (indirect_Loc - indirect_Cult)
#  3) Main effects (informationSource, scienceTeam) as marginal contrasts
#  4) Pairwise comparisons among the 4 cells
#
# Model: m_zoib_long_mixed_decision_point (brmsfit)
# Data:  data_long with factors informationSource, scienceTeam, decision_point (0/1), outcome decision in [0,1]

library(dplyr)
library(tidyr)
library(tidybayes)

# ---------- summary helper ----------
summ_contrast <- function(x, ci = 0.95) {
  x <- x[is.finite(x)]
  tibble(
    mean   = mean(x),
    sd     = sd(x),
    q_low  = unname(quantile(x, (1 - ci) / 2)),
    q_high = unname(quantile(x, 1 - (1 - ci) / 2)),
    p_gt0  = mean(x > 0),
    evid_ratio = mean(x > 0) / pmax(mean(x <= 0), .Machine$double.eps),
    n      = length(x)
  )
}

# ---------- core extractor ----------
epred_contrasts_2x2 <- function(fit,
                               data_long,
                               ndraws = 8000,
                               decision_point_value = 1L, # 0=decision1, 1=decision2
                               re_formula = NA,
                               ci = 0.95,
                               # explicit labels (must match levels(data_long$...))
                               info_levels = c("indirect", "direct"),
                               team_levels = c("Cultivation", "Localization")) {

  # 2x2 grid at decision point of interest
  nd <- expand_grid(
    informationSource = factor(info_levels, levels = levels(data_long$informationSource)),
    scienceTeam       = factor(team_levels, levels = levels(data_long$scienceTeam)),
    decision_point    = decision_point_value
  )

  # posterior expected predictions per draw, per cell
  cell_draws_long <- nd %>%
    add_epred_draws(
      fit,
      ndraws = ndraws,
      re_formula = re_formula,
      allow_new_levels = TRUE
    ) %>%
    group_by(.draw, informationSource, scienceTeam) %>%
    summarise(mu = mean(.epred, na.rm = TRUE), .groups = "drop") %>%
    mutate(cell = paste(informationSource, scienceTeam, sep = "_"))

  # wide: one column per cell
  cell_draws <- cell_draws_long %>%
    select(.draw, cell, mu) %>%
    pivot_wider(names_from = cell, values_from = mu)

  # expected column names
  nm_d_loc  <- "direct_Localization"
  nm_d_cult <- "direct_Cultivation"
  nm_i_loc  <- "indirect_Localization"
  nm_i_cult <- "indirect_Cultivation"

  needed <- c(nm_d_loc, nm_d_cult, nm_i_loc, nm_i_cult)
  missing <- setdiff(needed, names(cell_draws))
  if (length(missing) > 0) {
    stop(
      "Missing expected cell columns:\n",
      paste0("  - ", missing, collapse = "\n"),
      "\nAvailable columns:\n",
      paste0("  - ", names(cell_draws), collapse = "\n")
    )
  }

  # Delta interaction draws (your definition)
  delta_draws <- cell_draws %>%
    transmute(
      .draw,
      Delta_Team_byInfo = (`direct_Localization` - `direct_Cultivation`) -
                          (`indirect_Localization` - `indirect_Cultivation`)
    )

  # Main effects via marginalization (averaging over the other factor)
  # Main_Info: direct - indirect (averaged over team)
  # Main_Team: Localization - Cultivation (averaged over info)
  main_draws <- cell_draws %>%
    transmute(
      .draw,
      Main_Info = ((`direct_Localization` + `direct_Cultivation`)/2) -
                  ((`indirect_Localization` + `indirect_Cultivation`)/2),
      Main_Team = ((`direct_Localization` + `indirect_Localization`)/2) -
                  ((`direct_Cultivation` + `indirect_Cultivation`)/2)
    )

  # Pairwise among the 4 cells (reversed direction: b - a)
  cells_long2 <- cell_draws %>%
    select(.draw, all_of(needed)) %>%
    pivot_longer(-.draw, names_to = "cell", values_to = "mu")

  pairwise_draws <- cells_long2 %>%
    inner_join(cells_long2, by = ".draw", suffix = c("_a", "_b")) %>%
    filter(cell_a < cell_b) %>%
    mutate(
      contrast = paste0(cell_b, " - ", cell_a),  # reversed label
      value    = mu_b - mu_a                      # reversed direction
    ) %>%
    select(.draw, contrast, value)

  # Summaries
  summary_cells <- cell_draws_long %>%
    group_by(informationSource, scienceTeam) %>%
    summarise(
      mean   = mean(mu),
      sd     = sd(mu),
      q_low  = unname(quantile(mu, (1 - ci)/2)),
      q_high = unname(quantile(mu, 1 - (1 - ci)/2)),
      .groups = "drop"
    ) %>%
    mutate(effect = paste(informationSource, scienceTeam, sep = "_")) %>%
    select(effect, mean, sd, q_low, q_high)

  summary_delta <- summ_contrast(delta_draws$Delta_Team_byInfo, ci = ci) %>%
    mutate(effect = "Delta_Team_byInfo", .before = 1)

  summary_main <- bind_rows(
    summ_contrast(main_draws$Main_Info, ci = ci) %>%
      mutate(effect = "Main: informationSource (direct - indirect)", .before = 1),
    summ_contrast(main_draws$Main_Team, ci = ci) %>%
      mutate(effect = "Main: scienceTeam (Localization - Cultivation)", .before = 1)
  )

  summary_pairwise <- pairwise_draws %>%
    group_by(contrast) %>%
    summarise(
      mean   = mean(value),
      sd     = sd(value),
      q_low  = unname(quantile(value, (1 - ci)/2)),
      q_high = unname(quantile(value, 1 - (1 - ci)/2)),
      p_gt0  = mean(value > 0),
      evid_ratio = mean(value > 0) / pmax(mean(value <= 0), .Machine$double.eps),
      n      = n(),
      .groups = "drop"
    ) %>%
    arrange(desc(abs(mean))) %>%
    rename(effect = contrast)

  list(
    cell_draws = cell_draws,
    delta_draws = delta_draws,
    main_draws = main_draws,
    pairwise_draws = pairwise_draws,
    summary_cells = summary_cells,
    summary_delta = summary_delta,
    summary_main = summary_main,
    summary_pairwise = summary_pairwise
  )
}

# ---------- run on your model ----------
res <- epred_contrasts_2x2(
  fit = m_zoib_long_mixed_decision_point,
  data_long = data_long,
  ndraws = 8000,
  decision_point_value = 1L,  # decision2
  re_formula = NA,
  ci = 0.95
)

# 1) Posterior expected cell means (response scale)
res$summary_cells

# 2) Delta interaction (response scale)
res$summary_delta

# 3) Main effects (response scale)
res$summary_main

# 4) Pairwise comparisons (response scale; reversed direction: b - a)
res$summary_pairwise
```
```{r}
# ------------------------------------------------------------
# Posterior expected predictions for decision point 1 vs 2
# (marginalized over informationSource and scienceTeam)
# ------------------------------------------------------------
decision_point_effect_epred <- function(fit,
                                        data_long,
                                        ndraws = 8000,
                                        re_formula = NA,
                                        ci = 0.95) {

  # newdata grid: both decision points, all conditions
  nd <- expand_grid(
    informationSource = levels(data_long$informationSource),
    scienceTeam       = levels(data_long$scienceTeam),
    decision_point    = c(0L, 1L)  # 0 = decision1, 1 = decision2
  )

  draws <- nd %>%
    add_epred_draws(
      fit,
      ndraws = ndraws,
      re_formula = re_formula,
      allow_new_levels = TRUE
    ) %>%
    group_by(.draw, decision_point) %>%
    summarise(mu = mean(.epred, na.rm = TRUE), .groups = "drop") %>%
    mutate(dp = if_else(decision_point == 0L, "decision1", "decision2")) %>%
    select(.draw, dp, mu) %>%
    pivot_wider(names_from = dp, values_from = mu)

  # summaries
  s_decision1 <- summ_contrast(draws$decision1, ci) %>%
    mutate(effect = "E[belief] at decision 1", .before = 1)

  s_decision2 <- summ_contrast(draws$decision2, ci) %>%
    mutate(effect = "E[belief] at decision 2", .before = 1)

  s_diff <- summ_contrast(draws$decision2 - draws$decision1, ci) %>%
    mutate(effect = "Decision-point effect (decision2 − decision1)", .before = 1)

  bind_rows(s_decision1, s_decision2, s_diff)
}

# ------------------------------------------------------------
# Run
# ------------------------------------------------------------
dp_effects <- decision_point_effect_epred(
  fit = m_zoib_long_mixed_decision_point,
  data_long = data_long,
  ndraws = 8000,
  re_formula = NA
)

dp_effects
```


